---
title: "Prediciendo la Edad de los Abalones mediante Modelos Lineales"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# 1 Scope
The scope of this `R`-markdown notebook consists on predicting the age of abalones by means
of machine learning linear regression models.
In doing this, we use modern `R` syntax and libraries from `tidyverse` ecosystem such as `dplyR`, `ggplot` and `readr`

```{r, message=FALSE}
require(readr)
abalone <- readr::read_csv("datos/abalone.txt", col_names=F)
# Angel Muelas et al 
# Data file does not contain descriptor names
nombres_col <- c(
  "Sex",
  "Length",
  "Diameter",
  "Height",
  "Who_weight",
  "Shu_weight",
  "Viscera",
  "Shell",
  "Rings"
)
# Redefining tibble_df object
colnames(abalone) <- nombres_col
abalone$Sex <- as.factor(abalone$Sex)
```
# 2 EDA
Looking for null-values
```{r echo=TRUE, message=FALSE}
summary(abalone)
```
There are not null-values, however some nonsensical zero-values are present for 
the feature `Height`. Let us localize them

1. Classic syntax
```{r} 
# more familiar for pandas python users
abalone[abalone$Height==0,]
```
2. Modern syntax (`dplyr`)
```{r, message=FALSE}
require(dplyr)
dplyr::filter(abalone, Height==0)
```
Therefore, in order to filter null `Height` values we can proceed as follows
```{r}
abalone <- dplyr::filter(abalone, Height!=0)
```

## 2.1 Descriptive Analysis 
```{r}
# Equivalent to .describe()
summary(abalone)
```

```{r, message=FALSE}
# Qualitative cross-relations within ggpairs()
# including categorical variable: Sex
require(GGally)
ggpairs(
  data = abalone,
  mapping = ggplot2::aes(colour=Sex),
  diag = list(
    discrete="barDiag",                         # Facet BoxPlot
    continuous=wrap("densityDiag", alpha=0.5)   # density for numerical
  ),
  upper = list(continuous=wrap("cor", size=2)), # correlations on upper-diagonal
  lower = list(
    continuous=wrap("smooth", alpha=0.5)#,
  ),
progress = F # suppress verbosity
)
```

Excluding `Sex` feature (facet) in the discussion we have
```{r, message=FALSE}
ggpairs(
  data = abalone %>% select(-Sex),
  upper = list(continuous=wrap("cor", size=4)),
  lower = list(continuous=wrap("smooth", colour="blue")),
  progress = F
)
```
# 2.2 Outliers Detection

From visual inspection, clearly `Height` feature exhibits two outliers.
We can see it more precisely by using boxplot diagram restricted to this feature:
```{r, message=FALSE}
abalone %>% dplyr::select(Height) %>% boxplot()
```
In order to drop them, we can use filter
```{r, message=FALSE}
abalone <- 
  abalone %>% dplyr::filter(Height<0.5)
```
and test
```{r, message=FALSE}
abalone %>% select(Height) %>% boxplot()
```

Strong cross-correlations between features (all variables except \# Rings)
could be a sign of co linearity or multicolinearity (redundancy).

